Internal models of type theory and internal higher categories for the Univalent Foundations

### 2014-03-22 article-models-tt.tex
tweaking the positions of the QED signs
small changes

\begin{proof}
Let $\jfam{\Gamma}{A}$, $\jfam{\Gamma}{B}$, $\jfam{\Gamma}{C}$ and $\jhom{\Gamma}{A}{B}{f}$.
Then we have the judgmental equalities
\begin{align*}
\subst{f}{\ctxwk{A}{{B}{C}}}
& \jdeq 
  \subst{f}{\ctxwk{{A}{B}}{{A}{C}}}
  \\
& \jdeq 
  \ctxwk{A}{C}.
\end{align*}
\end{proof}

@@

\begin{proof}
Let $\jhom{\Gamma}{A}{B}{f}$, $\jhom{\Gamma}{B}{C}{g}$ and $\jfam{{\Gamma}{A}}{P}$.
Then we have the judgmental equalities
\begin{align*}
\ctxwk{P}{\jcomp{A}{f}{g}} 
& \jdeq 
  \ctxwk{P}{\subst{f}{\ctxwk{A}{g}}}
  \\
& \jdeq 
  \subst{\ctxwk{P}{f}}{\ctxwk{P}{{A}{g}}}
  \\
& \jdeq 
  \subst{\ctxwk{P}{f}}{\ctxwk{\ctxext{A}{P}}{g}}
  \\
& \jdeq 
  \jcomp{{A}{P}}{\ctxwk{P}{f}}{g}.
\end{align*}
Now let $\jterm{\Gamma}{B}{y}$ and $\jhom{\Gamma}{B}{C}{g}$. Then we have the
judgmental equalities
\begin{align*}
\jcomp{A}{\ctxwk{A}{y}}{g}
& \jdeq 
  \subst{\ctxwk{A}{y}}{\ctxwk{A}{g}}
  \\
& \jdeq 
  \ctxwk{A}{\subst{y}{g}}.
\end{align*}
For the third assertion, let $\jhom{\Gamma}{A}{B}{f}$ and $\jterm{\Gamma}{C}{z}$.
Then we have the judgmental equalities
\begin{align*}
\jcomp{A}{f}{\ctxwk{B}{z}} 
& \jdeq 
  \subst{f}{\ctxwk{A}{{B}{z}}}
  \\
& \jdeq 
  \subst{f}{\ctxwk{{A}{B}}{{A}{z}}}
  \\
& \jdeq 
  \ctxwk{A}{z}.
\end{align*}
\end{proof}

@@

\begin{proof}
Let $\jhom{{\Gamma}{A}}{P}{Q}{f}$, $\jhom{{\Gamma}{A}}{Q}{R}{g}$ and 
$\jterm{\Gamma}{A}{x}$.
Then we have the judgmental equalities
\begin{align*}
\subst{x}{\jcomp{A}{f}{g}}
& \jdeq 
  \subst{x}{{f}{\ctxwk{P}{g}}}
  \\
& \jdeq 
  \subst{{x}{f}}{{x}{\ctxwk{P}{g}}}
  \\
& \jdeq 
  \subst{{x}{f}}{\ctxwk{\subst{x}{P}}{\subst{x}{g}}}
  \\
& \jdeq 
  \jcomp{\subst{x}{P}}{\subst{x}{f}}{\subst{x}{g}}.
\end{align*}
Now let $\jhom{\Gamma}{A}{B}{f}$, $\jhom{\Gamma}{B}{C}{g}$ and $\jterm{\Gamma}{A}{x}$.
Then we have the judgmental equalities
\begin{align*}
\subst{x}{\jcomp{A}{f}{g}}
& \jdeq 
  \subst{x}{{f}{\ctxwk{A}{g}}}
  \\
& \jdeq 
  \subst{{x}{f}}{{x}{\ctxwk{A}{g}}}
  \\
& \jdeq 
  \subst{{x}{f}}{g}.
\end{align*}
\end{proof}

@@

\begin{proof}
Consider the families $\jfam{\Gamma}{A}$ and $\jfam{{\Gamma}{A}}{P}$. Then
we have the judgmental equalities
\begin{align*}
\tmext
  {\ctxwk{\ctxext{A}{P}}{A}}
  {\ctxwk{\ctxext{A}{P}}{P}}
  {\cprojfstf{A}{P}}
  {\cprojsndf{A}{P}}
& \jdeq 
  \unfold
  { \tmext
      {\ctxwk{\ctxext{A}{P}}{A}}
      {\ctxwk{\ctxext{A}{P}}{P}}
      {\cprojfstf{A}{P}}
      {\cprojsndf{A}{P}}
    }
  \\
& \jdeq 
  \subst
    { \idfunc[P]
      }
    { {\ctxwk{P}{\idfunc[A]}}
      {\idfunc[\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}]}
      }
  \\
& \jdeq 
  \subst
    { \idfunc[P]
      }
    { {\ctxwk{P}{\idfunc[A]}}
      {\ctxwk{\ctxext{A}{P}}{\idfunc[\ctxext{A}{P}]}}
      }
  \\
& \jdeq 
  \subst
    { \idfunc[P]
      }
    { {\ctxwk{P}{\idfunc[A]}}
      {\ctxwk{P}{{A}{\idfunc[\ctxext{A}{P}]}}}
      }
  \\
& \jdeq 
  \subst
    { \idfunc[P]
      }
    {\ctxwk
      {P}
      { \subst
        {\idfunc[A]}
        {\ctxwk{A}{\idfunc[\ctxext{A}{P}]}}
        }
      }
  \\
& \jdeq 
  \subst
    { \idfunc[A]
      }
    { \ctxwk{A}{\idfunc[\ctxext{A}{P}]}
      }
  \\
& \jdeq 
  \idfunc[\ctxext{A}{P}].
\end{align*}
\end{proof}

@@

\begin{proof}
To prove the first judgmental equality, note that
\begin{align*}
w 
& \jdeq 
  \subst{w}{\idfunc[\ctxext{A}{P}]} 
  \tag{by \autoref{idfunc-subst-defn}}\\
& \jdeq 
  \subst
    { w}
    { { \idfunc[P]}
      { { \ctxwk{P}{\idfunc[A]}}
        { \idfunc[\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}]}
        }
      }
  \tag{by \autoref{idfunc-ext-comp}}\\
& \jdeq 
  \subst
    { {w}
      {\idfunc[P]}
      }
    { {w}
      { { \ctxwk{P}{\idfunc[A]}
          }
        { \idfunc[\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}]
          }
        }
      }
  \tag{by \autoref{comp-ss-t}}\\
& \jdeq 
  \subst
    { {w}
      {\idfunc[P]}
      }
    { { {w}
        {\ctxwk{P}{\idfunc[A]}}
        }
      { {w}
        {\idfunc[\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}]}
        }
      }
  \tag{by \autoref{comp-ss-t}}\\
& \jdeq 
  \subst
    { {w}
      {\idfunc[P]}
      }
    { { {w}
        {\ctxwk{P}{\idfunc[A]}}
        }
      { {w}
        {\ctxwk{\ctxext{A}{P}}{\idfunc[\ctxext{A}{P}]}}
        }
      }
  \tag{by \autoref{idfunc-wk-comp}}\\
& \jdeq 
  \subst
    { {w}
      {\idfunc[P]}
      }
    { { {w}
        {\ctxwk{P}{\idfunc[A]}}
        }
      { \idfunc[\ctxext{A}{P}]
        }
      }
  \tag{by \autoref{defn-ws-4}}\\
& \jdeq 
  \tmext{A}{P}{\cprojfst{A}{P}{w}}{\cprojsnd{A}{P}{w}}
  \tag{by definition}
\end{align*}
To prove the second judgmental equality, let $\jterm{\Gamma}{A}{x}$ and
$\jterm{\Gamma}{\subst{x}{P}}{u}$. Then we have
\begin{align*}
\cprojfst{A}{P}{\tmext{A}{P}{x}{u}}
& \jdeq 
  \subst{\tmext{A}{P}{x}{u}}{\ctxwk{P}{\idfunc[A]}}
  \\
& \jdeq 
  \subst{u}{{x}{\ctxwk{P}{\idfunc[A]}}} 
  \\
& \jdeq 
  \subst{u}{\ctxwk{\subst{x}{P}}{\subst{x}{\idfunc[A]}}}
  \\
& \jdeq 
  \subst{x}{\idfunc[A]}
  \\
& \jdeq 
  x.
\end{align*}
To prove the third judgmental equality, note that
\begin{align*}
\cprojsnd{A}{P}{\tmext{A}{P}{x}{u}}
& \jdeq 
  \subst{\tmext{A}{P}{x}{u}}{\idfunc[P]}
  \\
& \jdeq 
  \subst{u}{{x}{\idfunc[P]}}
  \\
& \jdeq 
  \subst{u}{\idfunc[\subst{x}{P}]}
  \\
& \jdeq 
  u.
\end{align*}
\end{proof}

@@

\begin{proof}
Consider $\jterm{{\Gamma}{B}}{Q}{g}$ and $\jterm{{\Gamma}{B}}{\subst{g}{R}}{t}$.
Then we have the judgmental equalities
\begin{align*}
\ctxwk{A}{\ctxext{Q}{R}{g}{t}}
& \jdeq 
  \ctxwk{A}{\subst{t}{{g}{\idfunc[\ctxext{Q}{R}]}}}
  \\
& \jdeq 
  \subst{\ctxwk{A}{t}}{\ctxwk{A}{\subst{g}{\idfunc[\ctxext{Q}{R}]}}}
  \\
& \jdeq 
  \subst{\ctxwk{A}{t}}{{\ctxwk{A}{g}}{\ctxwk{A}{\idfunc[\ctxext{Q}{R}]}}}
  \\
& \jdeq 
  \subst{\ctxwk{A}{t}}{{\ctxwk{A}{g}}{\idfunc[\ctxwk{A}{\ctxext{Q}{R}}]}}
  \\
& \jdeq 
  \subst{\ctxwk{A}{t}}{{\ctxwk{A}{g}}{\idfunc[\ctxext{\ctxwk{A}{Q}}{\ctxwk{A}{R}}]}}
  \\
& \jdeq 
  \tmext{\ctxwk{A}{Q}}{\ctxwk{A}{R}}{\ctxwk{A}{g}}{\ctxwk{A}{t}}
\end{align*}
\end{proof}

@@

\begin{proof}
Consider $\jterm{{\Gamma}{B}}{Q}{g}$ and $\jterm{{\Gamma}{B}}{\subst{g}{R}}{t}$.
Then we have the judgmental equalities
\begin{align*}
\subst{x}{\tmext{Q}{R}{g}{t}}
& \jdeq 
  \subst{x}{{t}{{g}{\idfunc[\ctxext{Q}{R}]}}}
  \\
& \jdeq 
  \subst{{x}{t}}{{x}{{g}{\idfunc[\ctxext{Q}{R}]}}}
  \\
& \jdeq 
  \subst{{x}{t}}{{{x}{g}}{{x}{\idfunc[\ctxext{Q}{R}]}}}
  \\
& \jdeq 
  \subst{{x}{t}}{{{x}{g}}{\idfunc[\subst{x}{\ctxext{Q}{R}}]}}
  \\
& \jdeq 
  \subst{{x}{t}}{{{x}{g}}{\idfunc[\ctxext{\subst{x}{Q}}{\subst{x}{R}}]}}
  \\
& \jdeq 
  \tmext{\subst{x}{Q}}{\subst{x}{R}}{\subst{x}{g}}{\subst{x}{t}}.
\end{align*}
\end{proof}

@@

\begin{proof}
Let $\jhom{\Gamma}{A}{B}{f}$, $\jhom{\Gamma}{B}{C}{g}$, $\jfam{{\Gamma}{C}}{R}$
and $\jterm{{\Gamma}{B}}{\subst{g}{\ctxwk{B}{R}}}{w}$. Then we have the
judgmental equalities
\begin{align*}
\jcomp{A}{f}{\tmext{\ctxwk{B}{C}}{\ctxwk{B}{R}}{g}{w}}
& \jdeq 
  \subst{f}{\ctxwk{A}{\tmext{\ctxwk{B}{C}}{\ctxwk{B}{R}}{g}{w}}}
  \\
& \jdeq 
  \subst
    {f}
    {\tmext{\ctxwk{A}{{B}{C}}}{\ctxwk{A}{{B}{R}}}{\ctxwk{A}{g}}{\ctxwk{A}{w}}}
  \\
& \jdeq 
  \tmext
    {\ctxwk{A}{C}}
    {\ctxwk{A}{R}}
    {\subst{f}{\ctxwk{A}{g}}}
    {\subst{f}{\ctxwk{A}{w}}}
  \\
& \jdeq 
  \tmext{\ctxwk{A}{C}}{\ctxwk{A}{R}}{\jcomp{A}{f}{g}}{\jcomp{A}{f}{w}}.
\end{align*}
\end{proof}

>>>>

\begin{proof}
Let $\jfam{\Gamma}{A}$, $\jfam{\Gamma}{B}$, $\jfam{\Gamma}{C}$ and $\jhom{\Gamma}{A}{B}{f}$.
Then we have the judgmental equalities
\begin{align*}
\subst{f}{\ctxwk{A}{{B}{C}}}
& \jdeq 
  \subst{f}{\ctxwk{{A}{B}}{{A}{C}}}
  \\
& \jdeq 
  \ctxwk{A}{C}.
  \qedhere
\end{align*}
\end{proof}

@@

\begin{proof}
Let $\jhom{\Gamma}{A}{B}{f}$, $\jhom{\Gamma}{B}{C}{g}$ and $\jfam{{\Gamma}{A}}{P}$.
Then we have the judgmental equalities
\begin{align*}
\ctxwk{P}{\jcomp{A}{f}{g}} 
& \jdeq 
  \ctxwk{P}{\subst{f}{\ctxwk{A}{g}}}
  \\
& \jdeq 
  \subst{\ctxwk{P}{f}}{\ctxwk{P}{{A}{g}}}
  \\
& \jdeq 
  \subst{\ctxwk{P}{f}}{\ctxwk{\ctxext{A}{P}}{g}}
  \\
& \jdeq 
  \jcomp{{A}{P}}{\ctxwk{P}{f}}{g}.
\end{align*}
Now let $\jterm{\Gamma}{B}{y}$ and $\jhom{\Gamma}{B}{C}{g}$. Then we have the
judgmental equalities
\begin{align*}
\jcomp{A}{\ctxwk{A}{y}}{g}
& \jdeq 
  \subst{\ctxwk{A}{y}}{\ctxwk{A}{g}}
  \\
& \jdeq 
  \ctxwk{A}{\subst{y}{g}}.
\end{align*}
For the third assertion, let $\jhom{\Gamma}{A}{B}{f}$ and $\jterm{\Gamma}{C}{z}$.
Then we have the judgmental equalities
\begin{align*}
\jcomp{A}{f}{\ctxwk{B}{z}} 
& \jdeq 
  \subst{f}{\ctxwk{A}{{B}{z}}}
  \\
& \jdeq 
  \subst{f}{\ctxwk{{A}{B}}{{A}{z}}}
  \\
& \jdeq 
  \ctxwk{A}{z}.
  \qedhere
\end{align*}
\end{proof}

@@

\begin{proof}
Let $\jhom{{\Gamma}{A}}{P}{Q}{f}$, $\jhom{{\Gamma}{A}}{Q}{R}{g}$ and 
$\jterm{\Gamma}{A}{x}$.
Then we have the judgmental equalities
\begin{align*}
\subst{x}{\jcomp{A}{f}{g}}
& \jdeq 
  \subst{x}{{f}{\ctxwk{P}{g}}}
  \\
& \jdeq 
  \subst{{x}{f}}{{x}{\ctxwk{P}{g}}}
  \\
& \jdeq 
  \subst{{x}{f}}{\ctxwk{\subst{x}{P}}{\subst{x}{g}}}
  \\
& \jdeq 
  \jcomp{\subst{x}{P}}{\subst{x}{f}}{\subst{x}{g}}.
\end{align*}
Now let $\jhom{\Gamma}{A}{B}{f}$, $\jhom{\Gamma}{B}{C}{g}$ and $\jterm{\Gamma}{A}{x}$.
Then we have the judgmental equalities
\begin{align*}
\subst{x}{\jcomp{A}{f}{g}}
& \jdeq 
  \subst{x}{{f}{\ctxwk{A}{g}}}
  \\
& \jdeq 
  \subst{{x}{f}}{{x}{\ctxwk{A}{g}}}
  \\
& \jdeq 
  \subst{{x}{f}}{g}.
  \qedhere
\end{align*}
\end{proof}

@@

\begin{proof}
Consider the families $\jfam{\Gamma}{A}$ and $\jfam{{\Gamma}{A}}{P}$. Then
we have the judgmental equalities
\begin{align*}
\tmext
  {\ctxwk{\ctxext{A}{P}}{A}}
  {\ctxwk{\ctxext{A}{P}}{P}}
  {\cprojfstf{A}{P}}
  {\cprojsndf{A}{P}}
& \jdeq 
  \unfold
  { \tmext
      {\ctxwk{\ctxext{A}{P}}{A}}
      {\ctxwk{\ctxext{A}{P}}{P}}
      {\cprojfstf{A}{P}}
      {\cprojsndf{A}{P}}
    }
  \\
& \jdeq 
  \subst
    { \idfunc[P]
      }
    { {\ctxwk{P}{\idfunc[A]}}
      {\idfunc[\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}]}
      }
  \\
& \jdeq 
  \subst
    { \idfunc[P]
      }
    { {\ctxwk{P}{\idfunc[A]}}
      {\ctxwk{\ctxext{A}{P}}{\idfunc[\ctxext{A}{P}]}}
      }
  \\
& \jdeq 
  \subst
    { \idfunc[P]
      }
    { {\ctxwk{P}{\idfunc[A]}}
      {\ctxwk{P}{{A}{\idfunc[\ctxext{A}{P}]}}}
      }
  \\
& \jdeq 
  \subst
    { \idfunc[P]
      }
    {\ctxwk
      {P}
      { \subst
        {\idfunc[A]}
        {\ctxwk{A}{\idfunc[\ctxext{A}{P}]}}
        }
      }
  \\
& \jdeq 
  \subst
    { \idfunc[A]
      }
    { \ctxwk{A}{\idfunc[\ctxext{A}{P}]}
      }
  \\
& \jdeq 
  \idfunc[\ctxext{A}{P}].
  \qedhere
\end{align*}
\end{proof}

@@

\begin{proof}
To prove the first judgmental equality, note that
\begin{align*}
w 
& \jdeq 
  \subst{w}{\idfunc[\ctxext{A}{P}]} 
  \tag{by \autoref{idfunc-subst-defn}}\\
& \jdeq 
  \subst
    { w}
    { { \idfunc[P]}
      { { \ctxwk{P}{\idfunc[A]}}
        { \idfunc[\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}]}
        }
      }
  \tag{by \autoref{idfunc-ext-comp}}\\
& \jdeq 
  \subst
    { {w}
      {\idfunc[P]}
      }
    { {w}
      { { \ctxwk{P}{\idfunc[A]}
          }
        { \idfunc[\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}]
          }
        }
      }
  \tag{by \autoref{comp-ss-t}}\\
& \jdeq 
  \subst
    { {w}
      {\idfunc[P]}
      }
    { { {w}
        {\ctxwk{P}{\idfunc[A]}}
        }
      { {w}
        {\idfunc[\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}]}
        }
      }
  \tag{by \autoref{comp-ss-t}}\\
& \jdeq 
  \subst
    { {w}
      {\idfunc[P]}
      }
    { { {w}
        {\ctxwk{P}{\idfunc[A]}}
        }
      { {w}
        {\ctxwk{\ctxext{A}{P}}{\idfunc[\ctxext{A}{P}]}}
        }
      }
  \tag{by \autoref{idfunc-wk-comp}}\\
& \jdeq 
  \subst
    { {w}
      {\idfunc[P]}
      }
    { { {w}
        {\ctxwk{P}{\idfunc[A]}}
        }
      { \idfunc[\ctxext{A}{P}]
        }
      }
  \tag{by \autoref{defn-ws-4}}\\
& \jdeq 
  \tmext{A}{P}{\cprojfst{A}{P}{w}}{\cprojsnd{A}{P}{w}}
  \tag{by definition}
\end{align*}
To prove the second judgmental equality, let $\jterm{\Gamma}{A}{x}$ and
$\jterm{\Gamma}{\subst{x}{P}}{u}$. Then we have
\begin{align*}
\cprojfst{A}{P}{\tmext{A}{P}{x}{u}}
& \jdeq 
  \subst{\tmext{A}{P}{x}{u}}{\ctxwk{P}{\idfunc[A]}}
  \\
& \jdeq 
  \subst{u}{{x}{\ctxwk{P}{\idfunc[A]}}} 
  \\
& \jdeq 
  \subst{u}{\ctxwk{\subst{x}{P}}{\subst{x}{\idfunc[A]}}}
  \\
& \jdeq 
  \subst{x}{\idfunc[A]}
  \\
& \jdeq 
  x.
\end{align*}
To prove the third judgmental equality, note that
\begin{align*}
\cprojsnd{A}{P}{\tmext{A}{P}{x}{u}}
& \jdeq 
  \subst{\tmext{A}{P}{x}{u}}{\idfunc[P]}
  \\
& \jdeq 
  \subst{u}{{x}{\idfunc[P]}}
  \\
& \jdeq 
  \subst{u}{\idfunc[\subst{x}{P}]}
  \\
& \jdeq 
  u.
  \qedhere
\end{align*}
\end{proof}

@@

\begin{proof}
Consider $\jterm{{\Gamma}{B}}{Q}{g}$ and $\jterm{{\Gamma}{B}}{\subst{g}{R}}{t}$.
Then we have the judgmental equalities
\begin{align*}
\ctxwk{A}{\ctxext{Q}{R}{g}{t}}
& \jdeq 
  \ctxwk{A}{\subst{t}{{g}{\idfunc[\ctxext{Q}{R}]}}}
  \\
& \jdeq 
  \subst{\ctxwk{A}{t}}{\ctxwk{A}{\subst{g}{\idfunc[\ctxext{Q}{R}]}}}
  \\
& \jdeq 
  \subst{\ctxwk{A}{t}}{{\ctxwk{A}{g}}{\ctxwk{A}{\idfunc[\ctxext{Q}{R}]}}}
  \\
& \jdeq 
  \subst{\ctxwk{A}{t}}{{\ctxwk{A}{g}}{\idfunc[\ctxwk{A}{\ctxext{Q}{R}}]}}
  \\
& \jdeq 
  \subst{\ctxwk{A}{t}}{{\ctxwk{A}{g}}{\idfunc[\ctxext{\ctxwk{A}{Q}}{\ctxwk{A}{R}}]}}
  \\
& \jdeq 
  \tmext{\ctxwk{A}{Q}}{\ctxwk{A}{R}}{\ctxwk{A}{g}}{\ctxwk{A}{t}}
  \qedhere
\end{align*}
\end{proof}

@@

\begin{proof}
Consider $\jterm{{\Gamma}{B}}{Q}{g}$ and $\jterm{{\Gamma}{B}}{\subst{g}{R}}{t}$.
Then we have the judgmental equalities
\begin{align*}
\subst{x}{\tmext{Q}{R}{g}{t}}
& \jdeq 
  \subst{x}{{t}{{g}{\idfunc[\ctxext{Q}{R}]}}}
  \\
& \jdeq 
  \subst{{x}{t}}{{x}{{g}{\idfunc[\ctxext{Q}{R}]}}}
  \\
& \jdeq 
  \subst{{x}{t}}{{{x}{g}}{{x}{\idfunc[\ctxext{Q}{R}]}}}
  \\
& \jdeq 
  \subst{{x}{t}}{{{x}{g}}{\idfunc[\subst{x}{\ctxext{Q}{R}}]}}
  \\
& \jdeq 
  \subst{{x}{t}}{{{x}{g}}{\idfunc[\ctxext{\subst{x}{Q}}{\subst{x}{R}}]}}
  \\
& \jdeq 
  \tmext{\subst{x}{Q}}{\subst{x}{R}}{\subst{x}{g}}{\subst{x}{t}}.
  \qedhere
\end{align*}
\end{proof}

@@

\begin{proof}
Let $\jhom{\Gamma}{A}{B}{f}$, $\jhom{\Gamma}{B}{C}{g}$, $\jfam{{\Gamma}{C}}{R}$
and $\jterm{{\Gamma}{B}}{\subst{g}{\ctxwk{B}{R}}}{w}$. Then we have the
judgmental equalities
\begin{align*}
\jcomp{A}{f}{\tmext{\ctxwk{B}{C}}{\ctxwk{B}{R}}{g}{w}}
& \jdeq 
  \subst{f}{\ctxwk{A}{\tmext{\ctxwk{B}{C}}{\ctxwk{B}{R}}{g}{w}}}
  \\
& \jdeq 
  \subst
    {f}
    {\tmext{\ctxwk{A}{{B}{C}}}{\ctxwk{A}{{B}{R}}}{\ctxwk{A}{g}}{\ctxwk{A}{w}}}
  \\
& \jdeq 
  \tmext
    {\ctxwk{A}{C}}
    {\ctxwk{A}{R}}
    {\subst{f}{\ctxwk{A}{g}}}
    {\subst{f}{\ctxwk{A}{w}}}
  \\
& \jdeq 
  \tmext{\ctxwk{A}{C}}{\ctxwk{A}{R}}{\jcomp{A}{f}{g}}{\jcomp{A}{f}{w}}.
  \qedhere
\end{align*}
\end{proof}

\subsection{Judgmental trivial cofibrations}
\begin{defn}
Let $\jhom{\Gamma}{A}{B}{f}$ be a context morphism. We say that $f$ is a 
\emph{judgmental trivial cofibration} if the following inference rules are valid:
\begin{align*}
& \inference
  { \jfam{{\Gamma}{B}}{Q}
    \jterm{{\Gamma}{A}}{\jcomp{A}{f}{Q}}{g}
    }
  { \jterm{{\Gamma}{B}}{Q}{\jtcext{g}}
    }
& & \inference
  { \jfam{{\Gamma}{B}}{Q}
    \jterm{{\Gamma}{A}}{\jcomp{A}{f}{Q}}{g}
    }
  { \jtermeq{{\Gamma}{A}}{\jcomp{A}{f}{Q}}{\jcomp{A}{f}{\jtcext{g}}}{g}
    }
\end{align*}
\end{defn}

When $f$ is a context morphism from $A$ to $B$ in context $\Gamma$, 
finding a term of a family $Q$ over $\ctxext{\Gamma}{B}$ can be accomplished
by finding a term of the family $\jcomp{A}{f}{Q}$ over $\ctxext{\Gamma}{A}$.
Judgmental trivial cofibrations come with a computation rule. Thus, term
constructors of inductively defined types are going to be the major source of
examples of judgmental trivial cofibrations. Since identity functions behave
so nicely, they are judgmental trivial cofibrations too.

\begin{lem}
The identity function
$\jhom{\Gamma}{A}{A}{\idfunc[A]}$ is a judgmental trivial cofibration
for each family $A$ of contexts over $\Gamma$
\end{lem}

### 2014-03-22 article-models-ttconstr.tex
adding ideas to the introduction of section 4
small changes

\subsection{Dependent function types}

>>>>

\subsection{Dependent function types}
The dependent function type constructor bundles up the terms of a family $P$
of contexts over $\ctxext{\Gamma}{A}$ and produces a family $\mprd{A}{P}$ of
contexts over $\Gamma$. Its terms are copies of those of $P$. The operation
$\mprd{A}{\blank}$ also acts on terms of families -- this is called
$\lambda$-abstraction.

The dependent function type constructor not only acts on families $P$ of
contexts over $\ctxext{\Gamma}{A}$, but it also sends families $Q$ of contexts
over $\ctxext{{\Gamma}{A}}{P}$ to families $\mprd{A}{Q}$ over $\ctxext{\Gamma}
{\mprd{A}{P}}$. Moreover, we will also have a version of $\lambda$-abstraction
for terms of such families $Q$.

\begin{align*}
& \inference
    { \jfam{{\Gamma}{A}}{P}
      }
    { \jfam{\Gamma}{\mprd{A}{P}}
      }
& & \inference
      { \jfameq{\Gamma}{A}{A'}
        \jfameq{{\Gamma}{A}}{P}{P'}
        }
      { \jfameq{\Gamma}{\mprd{A}{P}}{\mprd{A'}{P'}}
        }
  \\
& \inference
    { \jterm{{\Gamma}{A}}{P}{f}
      }
    { \jterm{\Gamma}{\mprd{A}{P}}{\slam{A}{P}{f}}
      }
& & \inference
      { \jtermeq{{\Gamma}{A}}{P}{f}{f'}
        }
      { \jtermeq{\Gamma}{\mprd{A}{P}}{\slam{A}{P}{f}}{\slam{A}{P}{f'}}
        }
  \\
& \inference
    { \jterm{\Gamma}{\mprd{A}{P}}{g}
      }
    { \jterm{{\Gamma}{A}}{P}{\sev{g}}
      }
& & \inference
      { \jtermeq{\Gamma}{\mprd{A}{P}}{g}{g'}
        }
      { \jtermeq{{\Gamma}{A}}{P}{\sev{g}}{\sev{g'}}
        }
\end{align*}

\begin{defn}
Let $A$ and $B$ be families of contexts over $\Gamma$. Then we define
\begin{equation*}
\jfamdefn{\Gamma}{\jfun{A}{B}}{\sprd{A}{\ctxwk{A}{B}}}.
\end{equation*}
\end{defn}

\subsubsection{Dependent products are compatible with the empty context}
The empty context can appear in the domain and in the codomain of the dependent
function type constructor. We have the following inference rules explaining
what happens when the empty context appears in the domain:
\begin{align}
& \inference
    { \jfam{\Gamma}{P}
      }
    { \jfameq{\Gamma}{\mprd{\emptyf}{P}}{P}
      }
  \\
& \inference
    { \jterm{\Gamma}{P}{f}
      }
    { \jtermeq{\Gamma}{P}{\slam{A}{P}{f}}{f}
      }
  \\
& \inference
    { \jterm{\Gamma}{P}{g}
      }
    { \jtermeq{\Gamma}{P}{\sev{g}}{g}
      }
\end{align}
We have the following infernece rules explaining what happens when the empty
context appears in the codomain:
\begin{align}
& \inference
    { \jfam{\Gamma}{A}
      }
    { \jfameq{\Gamma}{\mprd{A}{\emptyf}}{\emptyf}
      }
\end{align}

\subsubsection{Dependent products are compatible with extension}
\begin{align}
& \inference
  { \jfam{{{{\Gamma}{A}}{B}}{P}}{Q}
    }
  { \jfameq
      {\Gamma}
      {\sprd{{A}{B}}{{P}{Q}}}
      {\ctxext{\sprd{A}{\sprd{B}{P}}}{\sprd{A}{\sprd{B}{Q}}}}
    }
  \\
& \inference
  { \jterm{{{\Gamma}{A}}{B}}{P}{f}
    }
  { \jtermeq
      {\Gamma}
      {\sprd{A}{\sprd{B}{P}}}
      {\slam{{A}{B}}{P}{f}}
      {\slam{A}{\sprd{B}{P}}{\slam{B}{P}{f}}}
    }
\end{align}
\subsubsection{Dependent products are compatible with weakening}
\subsubsection{Dependent products are compatible with substitution}?
\subsubsection{Dependent products are compatible with the identity functions}?
